k-Space Domain Parallel Transmit Pulse Design

### 2020-10-09 Manuscript_r1/response_to_reviews.tex

For a completely numerical paper however, I believe the stress testing should target not only what they have done, but also going to the extreme -- if they claim it is a problem they can solve.... Otherwise they should reduce their claim and state they can fix the problem partly by this proposed method.

For example, how does the proposed method handle the problem at the point, where the spatial domain method becomes infeasible.

I'm sure lots of pulse design will use the algorithm for other target regions (more detailed ones requiring higher resolution).

I couldn't quite find out what the 0.5-undersampling reduction factor meant in terms of pulse duration. If it translates to 666*2 time samples then again we are not near the 10,000 samples still. But maybe this should be described in more detail...
\\[0.2em]
\indent{\it \textcolor{blue}{The pulse design problem we evaluated in this work
did lead to very long compute times for the spatial domain designs and very large matrix sizes for an explicit pseudoinverse design,
so we believe the problem does represent a challenging but practical design scenario with which to evaluate the new method. 
At the same time, we have revised the max number of time points we cite in the Introduction from 10,000 down to 2,500, 
since even for the finest RF sampling time we know of on a major vendor scanner (4 us on GE), a 2,500 time point pulse is fairly long at 10 ms. 
We have added description to clarify how the SPINS trajectory was undersampled,
which was by reducing (or increasing, in the reduction factor 0.5 case) the number of cycles in each of the three trajectory segments.
We also now list the pulse durations for each reduction factor. 
To increase the paper's rigor we also added measured design times for the spatial domain method and the new k-space method
across reduction factors.}}
\\[1.2em]

@@

As discussed below, another reason to include mentions of even more pulse designs is that they tackle problems the proposed method does not. Similar to Hoyos-Idrobos [15] and Brunners [14] please discuss how the proposed method will handle if at all possible hard constraints. Is it possible?

I can only imagine that local SAR constraints will be quite comfortable to have with 24 channels. It is good that achieving the target profile is now faster with the proposed method, but how will the k-space domain method( and the spatial domain method for matter), handle local SAR? If you have 24 amplifiers tied together, are you all set to just rely on pulse (adaptive) regularisation or would you really desire to have hard peak and even average power constraints?

One can accept to only witness a comparison to the 2006 method, if the limitations of that and the proposed method are clearly labelled.

Yes, other methods might be much slower, but they handle problems you don't.
\\[0.2em]
\indent{\it \textcolor{blue}{We agree that we needed to tone down the language here. We have revised wording to indicate that nearly all parallel transmission studies have been formulated in the spatial domain, rather than being explicitly derived directly from the spatial domain method.
We have also added in the Discussion that it is not clear at this time whether or how hard power or SAR constraints could be directly built in to the method,
though the method would be compatible with parallel transmit VERSE methods.}}
\\[1.2em]

### 2020-10-09 Manuscript_r1/Introduction.tex
added k-traj figure to code, reduction factor descriptions

\par Despite the advantages of the spatial domain approach, 
it can have prohibitive computational requirements when the number of coils and the number of dimensions become large, 
for example in three-dimensional \cite{malik2012tailored,davids2016fast} or spectral-spatial pulse designs \cite{stenger2000three,setsompop2009,Malik:2010aa,yang2010four}, 
when there is a large number of coils \cite{orzada:2019},
and when off-resonance compensation is used.  
The spatial domain system matrix typically has dimensions $N_s \times (N_c \cdot N_t)$, 
where $N_s$ is the number of spatial and spectral locations defined in the target excitation pattern,
$N_c$ is the number of transmit channels, and $N_t$ is the number of time points in the pulse.
\textcolor{blue}{Typical values for $N_s$ range from 1,000 for 2D designs to 250,000 for 3D designs,
the number of coils varies from 8 to 32 on ultra-high field scanners,
and the number of time points varies from 500 to 2,500. 
Overall, for a 3D design the spatial domain system matrix can occupy upwards of 120 gigabytes of memory,
making it infeasible to construct and invert on most modern computers. } \revbox{R2.X}
Explicit matrix storage and inversion can be avoided using non-uniform fast Fourier transforms to evaluate matrix multiplies,
combined with iterative conjugate gradient solvers. 
In such designs, parallelization can be applied across coils or across time segments in off-resonance-compensated designs,
and GPU-based pulse designs have been described \cite{deng:2011}. 
However, parallelization cannot be applied across iterations;
hundreds of iterations are commonly used for 3D designs and thousands of iterations can be required for 
magnitude least-squares \cite{setsompop2008magnitude} 
or constrained pulse designs \cite{brunner2010optimal,hoyos:tmi:2014}. 
Furthermore, non-uniform fast Fourier transforms can require significant computation for the repeated gridding steps; 
in image reconstruction these can be eliminated since the Gram matrix is Toeplitz and can therefore be multiplied using fast Fourier transforms \cite{fessler2005toeplitz},
but this does not apply to parallel transmit pulse design since the Gram matrix is not Toeplitz. 
A further challenge for parallel transmit pulse design is that pulses must be computed for each subject,
as a pre-scan stage while the subject lies in the scanner.

>>>>

\par Despite the advantages of the spatial domain approach, 
it can have prohibitive computational requirements when the number of coils and the number of dimensions become large, 
for example in three-dimensional \cite{malik2012tailored,davids2016fast} or spectral-spatial pulse designs \cite{stenger2000three,setsompop2009,Malik:2010aa,yang2010four}, 
when there is a large number of coils \cite{orzada:2019},
and when off-resonance compensation is used.  
The spatial domain system matrix typically has dimensions $N_s \times (N_c \cdot N_t)$, 
where $N_s$ is the number of spatial and spectral locations defined in the target excitation pattern,
$N_c$ is the number of transmit channels, and $N_t$ is the number of time points in the pulse.
\textcolor{blue}{Typical values for $N_s$ range from 1,000 for 2D designs to 250,000 for 3D designs,
the number of coils varies from 8 to 32 on ultra-high field scanners,
and the number of time points varies from 500 to 2,500. 
Overall, for a 3D design the spatial domain system matrix can occupy upwards of 120 gigabytes of memory,
making it infeasible to construct and invert on most modern computers. } \revbox{R2.1}
Explicit matrix storage and inversion can be avoided using non-uniform fast Fourier transforms to evaluate matrix multiplies,
combined with iterative conjugate gradient solvers. 
In such designs, parallelization can be applied across coils or across time segments in off-resonance-compensated designs,
and GPU-based pulse designs have been described \cite{deng:2011}. 
However, parallelization cannot be applied across iterations;
hundreds of iterations are commonly used for 3D designs and thousands of iterations can be required for 
magnitude least-squares \cite{setsompop2008magnitude} 
or constrained pulse designs \cite{brunner2010optimal,hoyos:tmi:2014}. 
Furthermore, non-uniform fast Fourier transforms can require significant computation for the repeated gridding steps; 
in image reconstruction these can be eliminated since the Gram matrix is Toeplitz and can therefore be multiplied using fast Fourier transforms \cite{fessler2005toeplitz},
but this does not apply to parallel transmit pulse design since the Gram matrix is not Toeplitz. 
A further challenge for parallel transmit pulse design is that pulses must be computed for each subject,
as a pre-scan stage while the subject lies in the scanner.

### 2020-10-09 Manuscript_r1/Method.tex
added k-traj figure to code, reduction factor descriptions

\subsection*{Excitation k-Space Undersampling}
An important application of parallel transmission is the reduction of multidimensional pulse durations by excitation k-space undersampling.
To compare the spatial domain and k-space domain methods across undersampling factors,
the number of polar and azimuthal rotations were increased and reduced relative to the reference design described above, 
for reduction factors of 0.5, 2, and 4, 
\textcolor{blue}{where the factor 0.5 trajectory comprised X additional segments for approximately 2$\times$-denser sampling of excitation k-space.} \revbox{2.X}

>>>>

\subsection*{Excitation k-Space Undersampling}
An important application of parallel transmission is the reduction of multidimensional pulse durations by excitation k-space undersampling.
\textcolor{blue}{To compare the spatial domain and k-space domain methods across undersampling factors,
the number of polar and azimuthal rotations in each segment of the SPINS trajectory were increased and reduced relative to the reference design described above, 
by factors of 0.5, 2, and 4. 
That is, for a reduction factor of 0.5, the number of polar and azimuthal rotations in the trajectory were doubled for each of the three segments,
and they were halved for a reduction factor of 2.
This yielded pulse durations of 20.5 ms (reduction factor 0.5), 5.3 ms (reduction factor 2), and 2.6 ms (reduction factor 4).} \revbox{R2.1}
